---
title: "Analysis of Crowd Source Data"
output: html_document
---

# Load needed libraries
```{r eval = FALSE, echo = FALSE}
# Run this section manually only once to import all the dependencies into RStudio
install.packages("pROC")
install.packages("randomForest")
install.packages("survival")
install.packages("lattice")
install.packages("gbm")
install.packages("foreach")
install.packages("gam")
install.packages("Matrix")
install.packages("glmnet")
install.packages("tree")
```

```{r}
library(pROC)
library(randomForest)
library(survival)
library(lattice)
# library(splines) # loaded by gam as a dependency
# library(foreach) # loaded by gam as a dependency
library(parallel)
library(gbm)
library(gam)
library(Matrix)
library(glmnet)
library(tree)
```

# Load data
```{r}
crowd = read.csv("crowd.csv")
```

# Assign test set
```{r}
set.seed(12345)
train = runif(nrow(crowd))<.7   # pick train/test split
table(train)
names(crowd)
summary(crowd)
```

# Examine correlation matrix
```{r}
round(cor(crowd), 2)
```

# Ridge (Contributor+Content)
```{r}
x = model.matrix(y~.-pastideas-votes-comments, crowd)
fit.ridge = glmnet(x[train,], crowd$y[train], family="binomial", alpha=0)
fit.cv.ridge = cv.glmnet(x[train,], crowd$y[train], family="binomial", alpha=0)
plot(fit.ridge, xvar = "lambda")
abline(v=log(fit.cv.ridge$lambda.min))
```

```{r}
predict(fit.ridge, x[!train,], s=fit.cv.ridge$lambda.min, type="coef")
```

```{r}
phat = predict(fit.ridge, x[!train,], s=fit.cv.ridge$lambda.min, type="response")
auc(crowd$y[!train], as.vector(phat))
```
<!--
[1] 0.6252981
-->

# Ridge (All variables)
```{r}
x = model.matrix(y~., crowd)
fit.ridge = glmnet(x[train,], crowd$y[train], family="binomial", alpha=0)
fit.cv.ridge = cv.glmnet(x[train,], crowd$y[train], family="binomial", alpha=0)
plot(fit.ridge, xvar = "lambda")
abline(v=log(fit.cv.ridge$lambda.min))
```

```{r}
predict(fit.ridge, x[!train,], s=fit.cv.ridge$lambda.min, type="coef")
```

```{r}
phat = predict(fit.ridge, x[!train,], s=fit.cv.ridge$lambda.min, type="response")
auc(crowd$y[!train], as.vector(phat))
```

# Lasso (Contributor+Content)
```{r}
x = model.matrix(y~.-pastideas-votes-comments, crowd)
fit.lasso = glmnet(x[train,], crowd$y[train], family="binomial", alpha=1)
fit.cv.lasso = cv.glmnet(x[train,], crowd$y[train], family="binomial", alpha=1)
plot(fit.lasso, xvar = "lambda")
abline(v=log(fit.cv.lasso$lambda.min))
```

```{r}
predict(fit.lasso, x[!train,], s=fit.cv.lasso$lambda.min, type="coef")
```

```{r}
phat = predict(fit.lasso, x[!train,], s=fit.cv.lasso$lambda.min, type="response")
auc(crowd$y[!train], as.vector(phat))
```

# Lasso (All variables)
```{r}
x = model.matrix(y~., crowd)
fit.lasso = glmnet(x[train,], crowd$y[train], family="binomial", alpha=1)
fit.cv.lasso = cv.glmnet(x[train,], crowd$y[train], family="binomial", alpha=1)
plot(fit.lasso, xvar = "lambda")
abline(v=log(fit.cv.lasso$lambda.min))
```

```{r}
predict(fit.lasso, x[!train,], s=fit.cv.lasso$lambda.min, type="coef")
```

```{r}
phat = predict(fit.lasso, x[!train,], s=fit.cv.lasso$lambda.min, type="response")
auc(crowd$y[!train], as.vector(phat)) # test set AUC = 0.7916172 
```

# GAM (Contributor+Content)
```{r}
fit.gam = gam(y ~ s(pastaccept)+s(commentsC)+s(X1)+s(X2)+s(X3)+s(X4)+s(X5)+s(X6)+s(X7)+s(X8)+s(X9)+s(X10)+s(X11)+s(age)+s(month)+s(diversity), binomial, crowd[train,])
summary(fit.gam)
```

```{r}
par(mfrow=c(1,3))
plot(fit.gam, se=T)
```

```{r}
par(mfrow=c(1,1))
phat = predict(fit.gam, crowd[!train,], type="response")
auc(crowd$y[!train], phat) 
```

# GAM (All variables)
```{r}
fit.gam = gam(y ~ ., binomial, crowd[train,])
summary(fit.gam)
```

```{r}
par(mfrow=c(1,3))
plot(fit.gam, se=T)
```

```{r}
par(mfrow=c(1,1))
phat = predict(fit.gam, crowd[!train,], type="response")
auc(crowd$y[!train], phat) 
```

# Tree (Contributor+Content)
```{r}
fit.tree = tree(y~.-pastideas-votes-comments, data=crowd[train,])
fit.tree
```

```{r}
plot(fit.tree, type="uniform")
text(fit.tree)
plot(cv.tree(fit.tree))
```

```{r}
phat = predict(fit.tree, crowd[!train,])
auc(crowd$y[!train], phat)
```

# Tree (All variables)
```{r}
fit.tree = tree(y~., data=crowd[train,])
fit.tree
```

```{r}
plot(fit.tree, type="uniform")
text(fit.tree)
plot(cv.tree(fit.tree))
```

```{r}
phat = predict(fit.tree, crowd[!train,])
auc(crowd$y[!train], phat)
```

# Random Forest (Contributor+Content)
```{r}
set.seed(12345)
fit = randomForest(x=crowd[train, c(1,2,4,5,6,9:18)], y=factor(1-crowd$y[train]), xtest=crowd[!train,c(1,2,4,5,6,9:18)], ntree=100, keep.forest=T)
fit
```

```{r}
hist(fit$test$votes[,2], main="Predicted probabilities for test set")
```

```{r}
varImpPlot(fit)
```

```{r}
partialPlot(fit, crowd[train,], "diversity")
partialPlot(fit, crowd[train,], "X8")
partialPlot(fit, crowd[train,], "X3")
```

```{r}
auc(1-crowd$y[!train], fit$test$votes[,2])
```

```{r}
plot.roc(crowd$y[!train], fit$test$votes[,2], print.auc=T)
```

# Random Forest (All variables)
```{r}
set.seed(12345)
fit = randomForest(x=crowd[train,], y=factor(1-crowd$y[train]), xtest=crowd[!train,], ntree=100, keep.forest=T)
fit
```

```{r}
auc(1-crowd$y[!train], fit$test$votes[,2])
```

```{r}
plot.roc(crowd$y[!train], fit$test$votes[,2], print.auc=T)
```

# GBM (Contributor+Content)
```{r}
fit.gbm = gbm(y~.-pastideas-votes-comments, data=crowd[train,], n.trees=1000)
fit.gbm
summary(fit.gbm)
```

```{r}
phat = predict(fit.gbm, crowd[!train,], n.trees=1000, type="response")
auc(crowd$y[!train], phat)
phat = predict(fit.gbm, crowd[!train,], n.trees=100, type="response")
auc(crowd$y[!train], phat)
phat = predict(fit.gbm, crowd[!train,], n.trees=500, type="response")
auc(crowd$y[!train], phat)
```

# GBM (All variables)
```{r}
fit.gbm = gbm(y~., data=crowd[train,], n.trees=1000)
fit.gbm
summary(fit.gbm)
```

```{r}
phat = predict(fit.gbm, crowd[!train,], n.trees=1000, type="response")
auc(crowd$y[!train], phat)
phat = predict(fit.gbm, crowd[!train,], n.trees=100, type="response")
auc(crowd$y[!train], phat)
phat = predict(fit.gbm, crowd[!train,], n.trees=500, type="response")
auc(crowd$y[!train], phat)
phat = predict(fit.gbm, crowd[!train,], n.trees=700, type="response")
auc(crowd$y[!train], phat)
```

# Logistic Regression (Contributor+Content)
```{r}
fit = glm(y~.-pastideas-votes-comments, binomial, crowd[train,])
log_odd_hat = predict(fit, data.frame(crowd[!train,]))
phat = 1/(1+exp(-log_odd_hat))
auc(crowd$y[!train], as.vector(phat))
```

# Logistic Regression (All variables)
```{r}
fit = glm(y~., binomial, crowd[train, ])
log_odd_hat = predict(fit, data.frame(crowd[!train,]))
phat = 1/(1+exp(-log_odd_hat))
auc(crowd$y[!train], as.vector(phat))
```